---
title: 'Visual Story Telling Part 1: Green Building'
author: "Carissa Ing"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(factoextra)
library(cluster)
```

```{r, message=FALSE}
buildings <- read_csv("greenbuildings.csv") 
```

## Clean Data

```{r}
b <- buildings %>% 
  drop_na() %>%    
  filter(leasing_rate >= 0.1) %>% 
  select(-CS_PropertyID)
```

```{r}
green <- b %>% filter(green_rating > 0)

not_green <- b %>% filter(green_rating < 1)
```


## Assumption 1: The square footage of a green building is approximately the same as that of a non-green building

```{r}
print("Avg Rentable Sq Ft of Green Buildings:")
mean(green$size)

print("Avg Rentable Sq Ft of Non-Green Buildings:")
mean(not_green$size)
```

The average square footage available for rent in green buildings is 1.4 times larger than in non-green buildings. If both were to achieve 100% occupancy, a green building would generate 1.4 times more rental income per square foot than a non-green building. This supports the guru's claim that the construction would be profitable. However, it is highly unlikely for a building to be fully occupied in its first few years.


## Assumption 2: Green buildings are fully occupied, quickly.

```{r, message=FALSE, warning=FALSE}
green %>% 
  filter(age < 20) %>% 
  mutate(age = as.factor(age)) %>% 
  group_by(age) %>% 
  summarise(avg_leasing_rate = mean(leasing_rate)) %>% 
  ggplot(aes(x = age, y = avg_leasing_rate)) +
  geom_point(size = 2) +
  ylim(0, 100) +
  labs(
    title = "Avg Leasing Rate As Green Buildings Age",
    x = "Age of Building (Yrs)",
    y = "AVg Building Occupany %"
  )
```

The figure above shows that green buildings, on average, do not achieve over 90% occupancy until about their third year. The stats guru claimed it would take at most 9 years to recoup the cost of development and begin making a profit based on a 90% occupancy rate. However, that timeline is likely to be extended by about 2 years. While a 2-year extension is not significant given the typically long lifespan of green buildings, it is still only an estimate, assuming green buildings earn $2.60 more per square foot than non-green buildings, as the stats guru suggested. I will be clustering to find a more accurate estimate of this rent difference.


## Clustering for Rent Difference

```{r}
selected_features <- b %>%
  select(Rent, leasing_rate, size, stories, age, green_rating, cluster_rent, Precipitation, Gas_Costs, 
         Electricity_Costs)

selected_features <- na.omit(selected_features)

scaled_features <- scale(selected_features)
```


### Dimensionality Reduction

```{r}
# Perform PCA
pca_result <- prcomp(scaled_features, scale. = TRUE)

#Choose num components based on >90% variance explained
fviz_eig(pca_result)
selected_pca <- pca_result$x[, 1:8]

```

### K-Means Clustering

```{r}
fviz_nbclust(selected_pca, kmeans, method = "wss")

set.seed(85)
kmeans_result <- kmeans(selected_pca, centers = 6, nstart = 25)

b$kmeans_cluster <- kmeans_result$cluster
#fviz_cluster(kmeans_result, data = selected_pca)
```

```{r, message=FALSE}
green <- b %>% filter(green_rating >0)

not_green <- b %>% filter(green_rating < 1)
```

```{r}
# See "centroids" of each cluster
ng_cluster_summary <- not_green %>%
  group_by(kmeans_cluster) %>%
  summarize(med_rent_ng = median(Rent),
            med_size_ng = median(size),
            med_stories_ng = median(stories))
            

g_cluster_summary <- green %>%
  filter(green_rating == 1) %>%
  group_by(kmeans_cluster) %>%
  summarize(med_rent_g = median(Rent),
            med_size_g = median(size),
            med_stories_g = median(stories))

combined_summary <- merge(g_cluster_summary, ng_cluster_summary, by = "kmeans_cluster", all = TRUE)

combined_summary <- combined_summary %>%
  select(kmeans_cluster, sort(names(combined_summary)[names(combined_summary) != "kmeans_cluster"]))

print(combined_summary)
```

The table above shows the median values for each cluster of green buildings (denoted by the suffix "_g") and non-green buildings (denoted by the suffix "_ng"). Missing values indicate that there are no green or non-green buildings within that cluster. For instance, clusters 1 and 2 group only non-green buildings, while cluster 5 contains only green buildings. Clusters 3, 4, and 6 are less effective at distinguishing between the two types.

The planned development is a 15-story, 250,000 sq ft building, which will likely fall into cluster 5. Thus, the estimated future rent price will be \$27.68 per sq ft.

Between clusters 1 and 2, which focus on non-green buildings, cluster 2 has median size and story values most similar to those in cluster 5 and will serve as the standard of comparison. Cluster 2 has a median rent price of \$19.50 per sq ft, meaning that green buildings are expected to command about \$8.18 more per sq ft.

The total additional profit per year from building a green building rather than a non-green one is \$8.18 × 250,000 sq ft = \$2,045,000. To recoup the cost of development, it would take approximately \$5,000,000 / \$2,045,000 = 2.44 years. Given this relatively short time span compared to the long lifespan of green buildings, I agree with the stat guru that proceeding with the green building is a profitable decision.
